---
title: "MA317 Final Assignment"
author: "Jasper Yang"
date: "12/6/2022"
output:
  word_document: default
  html_document:
    df_print: paged
---

```{r setup, include= FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Read in the data
LE_data_raw <- read.csv("Life_Expectancy_Data1.csv")
LE_data <- subset(LE_data_raw, select = -c(Country.Code,EG.FEC.RNEW.ZS)) #we deleted .. Renewable energy consumption because it had 100% missingness
LE_data$SP.POP.TOTL <- LE_data$SP.POP.TOTL/100000 # Change units of population to be more interpretable
LE_data$SH.HIV.INCD <- LE_data$SH.HIV.INCD/1000 # Change units of HIV adults to 1000s

### load packages
library(ISLR)
library(tidyr)
library(dplyr)
library(ggplot2)
library(corrplot)
library(flextable)
library(mice)
library(faraway)
library(cowplot)
library(ggthemes)
library(ggsci)
library(miceadds)
library(officer)
library(multcomp)
library(car)

# options
m <- 10 # number of imputations
```

## Question 1

1. Exploratory Data Analysis: describe the data and deal with missing values
(a) Analyse using descriptive statistics (both graphical and numerical representations) and R the Life Expectancy data1.csv dataset. [12 marks]

```{r, echo = FALSE}
# N obs
dim(LE_data) # 217 observations, 29 columns

# N missing for LE
table(is.na(LE_data$SP.DYN.LE00.IN))
table(is.na(LE_data$SP.DYN.LE00.IN))/nrow(LE_data)


# Max and min
max(LE_data$SP.DYN.LE00.IN, na.rm = TRUE)
dplyr::arrange(LE_data, desc(SP.DYN.LE00.IN))$Country.Name[1]

min(LE_data$SP.DYN.LE00.IN, na.rm = TRUE)
dplyr::arrange(LE_data, SP.DYN.LE00.IN)$Country.Name[1]


vec <- c()
for (i in 1:ncol(LE_data)){
  vec[i] <- class(LE_data[,1])
}

table(vec)



# Correlation plot
#LE_cor <- cor(LE_data)
#corrplot.mixed(LE_cor, lower.col = "black", number.cex = .7)

```

```{r, echo = FALSE}
# Names key:
new_names <- c("Country.Name", "Continent",
               "life_expectancy", "electricity_access",
               "income_growth", "income_growth_pc",
               "children_HIV", "children_out_of_school",
               "ed_attain_primary_rate", "ed_attain_bach_rate",
               "infant_mortality_rate", "primary_completion_rate",
               "literacy_rate", "real_interest_rate",
               "pop_growth_perc_annual", "pop_dens",
               "pop_total", "health_expend_pc",
               "health_expend_perc_gdp", "unemployment_rate",
               "gdp_growth", "gdp_pc", "birth_rate",
               "adults_HIV",
               "safe_drinking_water", "poverty_ratio",
               "ed_duration")

names(LE_data) <- new_names
```

## Table by Continents
```{r, echo = FALSE, include = FALSE}
## Split data by continent for table
LE_Africa <- dplyr::filter(LE_data, Continent == 'Africa')
LE_Asia <- dplyr::filter(LE_data, Continent == 'Asia') 
LE_Oceania <- dplyr::filter(LE_data, Continent == 'Australia/Oceania')
LE_Europe <- dplyr::filter(LE_data, Continent == 'Europe')
LE_NA <- dplyr::filter(LE_data, Continent == 'North America')
LE_SA <- dplyr::filter(LE_data, Continent == 'South America')

### Choose a variable for each important thing

## Life Expectancy
life_expectancy_Africa <- paste0(
  sprintf("%.1f",mean(LE_Africa$life_expectancy, na.rm = TRUE)), ' (', sprintf("%.1f", sd(LE_Africa$life_expectancy, na.rm = TRUE)), ")")
life_expectancy_Asia <- paste0(
  sprintf("%.1f",mean(LE_Asia$life_expectancy, na.rm = TRUE)), ' (', sprintf("%.1f",sd(LE_Asia$life_expectancy, na.rm = TRUE)), ")")
life_expectancy_Oceania <- paste0(
  sprintf("%.1f",mean(LE_Oceania$life_expectancy, na.rm = TRUE)), ' (', sprintf("%.1f",sd(LE_Oceania$life_expectancy, na.rm = TRUE)), ")")
life_expectancy_Europe <- paste0(
  sprintf("%.1f",mean(LE_Europe$life_expectancy, na.rm = TRUE)), ' (', sprintf("%.1f",sd(LE_Europe$life_expectancy, na.rm = TRUE)), ")")
life_expectancy_NA <- paste0(
  sprintf("%.1f",mean(LE_NA$life_expectancy, na.rm = TRUE)), ' (', sprintf("%.1f",sd(LE_NA$life_expectancy, na.rm = TRUE)), ")")
life_expectancy_SA <- paste0(
  sprintf("%.1f",mean(LE_SA$life_expectancy, na.rm = TRUE)), ' (', sprintf("%.1f",sd(LE_SA$life_expectancy, na.rm = TRUE)), ")")
life_expectancy_All <- paste0(
  sprintf("%.1f",mean(LE_data$life_expectancy, na.rm = TRUE)), ' (', sprintf("%.1f",sd(LE_data$life_expectancy, na.rm = TRUE)), ")")

table1_life_expectancy <- c(life_expectancy_Africa,
                        life_expectancy_Asia,
                        life_expectancy_Oceania, 
                        life_expectancy_Europe,
                        life_expectancy_NA,
                        life_expectancy_SA,
                        life_expectancy_All)

## Primary Educational Attainment (% 25+) 
primary_completion_Africa <- paste0(
  sprintf("%.1f",mean(LE_Africa$primary_completion_rate, na.rm = TRUE)), ' (', sprintf("%.1f", sd(LE_Africa$primary_completion_rate, na.rm = TRUE)), ")")
primary_completion_Asia <- paste0(
  sprintf("%.1f",mean(LE_Asia$primary_completion_rate, na.rm = TRUE)), ' (', sprintf("%.1f",sd(LE_Asia$primary_completion_rate, na.rm = TRUE)), ")")
primary_completion_Oceania <- paste0(
  sprintf("%.1f",mean(LE_Oceania$primary_completion_rate, na.rm = TRUE)), ' (', sprintf("%.1f",sd(LE_Oceania$primary_completion_rate, na.rm = TRUE)), ")")
primary_completion_Europe <- paste0(
  sprintf("%.1f",mean(LE_Europe$primary_completion_rate, na.rm = TRUE)), ' (', sprintf("%.1f",sd(LE_Europe$primary_completion_rate, na.rm = TRUE)), ")")
primary_completion_NA <- paste0(
  sprintf("%.1f",mean(LE_NA$primary_completion_rate, na.rm = TRUE)), ' (', sprintf("%.1f",sd(LE_NA$primary_completion_rate, na.rm = TRUE)), ")")
primary_completion_SA <- paste0(
  sprintf("%.1f",mean(LE_SA$primary_completion_rate, na.rm = TRUE)), ' (', sprintf("%.1f",sd(LE_SA$primary_completion_rate, na.rm = TRUE)), ")")
primary_completion_All <- paste0(
  sprintf("%.1f",mean(LE_data$primary_completion_rate, na.rm = TRUE)), ' (', sprintf("%.1f",sd(LE_data$primary_completion_rate, na.rm = TRUE)), ")")

table1_primary_completion <- c(primary_completion_Africa,
                        primary_completion_Asia,
                        primary_completion_Oceania, 
                        primary_completion_Europe,
                        primary_completion_NA,
                        primary_completion_SA,
                        primary_completion_All)

## Population Density
pop_density_Africa <- paste0(
  sprintf("%.1f",mean(LE_Africa$pop_dens, na.rm = TRUE)), ' (', sprintf("%.1f", sd(LE_Africa$pop_dens, na.rm = TRUE)), ")")
pop_density_Asia <- paste0(
  sprintf("%.1f",mean(LE_Asia$pop_dens, na.rm = TRUE)), ' (', sprintf("%.1f",sd(LE_Asia$pop_dens, na.rm = TRUE)), ")")
pop_density_Oceania <- paste0(
  sprintf("%.1f",mean(LE_Oceania$pop_dens, na.rm = TRUE)), ' (', sprintf("%.1f",sd(LE_Oceania$pop_dens, na.rm = TRUE)), ")")
pop_density_Europe <- paste0(
  sprintf("%.1f",mean(LE_Europe$pop_dens, na.rm = TRUE)), ' (', sprintf("%.1f",sd(LE_Europe$pop_dens, na.rm = TRUE)), ")")
pop_density_NA <- paste0(
  sprintf("%.1f",mean(LE_NA$pop_dens, na.rm = TRUE)), ' (', sprintf("%.1f",sd(LE_NA$pop_dens, na.rm = TRUE)), ")")
pop_density_SA <- paste0(
  sprintf("%.1f",mean(LE_SA$pop_dens, na.rm = TRUE)), ' (', sprintf("%.1f",sd(LE_SA$pop_dens, na.rm = TRUE)), ")")
pop_density_All <- paste0(
  sprintf("%.1f",mean(LE_data$pop_dens, na.rm = TRUE)), ' (', sprintf("%.1f",sd(LE_data$pop_dens, na.rm = TRUE)), ")")

table1_pop_density <- c(pop_density_Africa,
                        pop_density_Asia,
                        pop_density_Oceania, 
                        pop_density_Europe,
                        pop_density_NA,
                        pop_density_SA,
                        pop_density_All)

## Health Expenditure per Capita
health_expenditure_Africa <- paste0(
  sprintf("%.1f",mean(LE_Africa$health_expend_pc, na.rm = TRUE)), ' (', sprintf("%.1f", sd(LE_Africa$health_expend_pc, na.rm = TRUE)), ")")
health_expenditure_Asia <- paste0(
  sprintf("%.1f",mean(LE_Asia$health_expend_pc, na.rm = TRUE)), ' (', sprintf("%.1f",sd(LE_Asia$health_expend_pc, na.rm = TRUE)), ")")
health_expenditure_Oceania <- paste0(
  sprintf("%.1f",mean(LE_Oceania$health_expend_pc, na.rm = TRUE)), ' (', sprintf("%.1f",sd(LE_Oceania$health_expend_pc, na.rm = TRUE)), ")")
health_expenditure_Europe <- paste0(
  sprintf("%.1f",mean(LE_Europe$health_expend_pc, na.rm = TRUE)), ' (', sprintf("%.1f",sd(LE_Europe$health_expend_pc, na.rm = TRUE)), ")")
health_expenditure_NA <- paste0(
  sprintf("%.1f",mean(LE_NA$health_expend_pc, na.rm = TRUE)), ' (', sprintf("%.1f",sd(LE_NA$health_expend_pc, na.rm = TRUE)), ")")
health_expenditure_SA <- paste0(
  sprintf("%.1f",mean(LE_SA$health_expend_pc, na.rm = TRUE)), ' (', sprintf("%.1f",sd(LE_SA$health_expend_pc, na.rm = TRUE)), ")")
health_expenditure_All <- paste0(
  sprintf("%.1f",mean(LE_data$health_expend_pc, na.rm = TRUE)), ' (', sprintf("%.1f",sd(LE_data$health_expend_pc, na.rm = TRUE)), ")")

table1_health_expenditure <- c(health_expenditure_Africa, 
                               health_expenditure_Asia, 
                               health_expenditure_Oceania, 
                               health_expenditure_Europe,
                               health_expenditure_NA,
                               health_expenditure_SA, 
                               health_expenditure_All)

## GDP per Capita
GDP_Africa <- paste0(
  sprintf("%.1f",mean(LE_Africa$gdp_pc, na.rm = TRUE)), ' (', sprintf("%.1f", sd(LE_Africa$gdp_pc, na.rm = TRUE)), ")")
GDP_Asia <- paste0(
  sprintf("%.1f",mean(LE_Asia$gdp_pc, na.rm = TRUE)), ' (', sprintf("%.1f",sd(LE_Asia$gdp_pc, na.rm = TRUE)), ")")
GDP_Oceania <- paste0(
  sprintf("%.1f",mean(LE_Oceania$gdp_pc, na.rm = TRUE)), ' (', sprintf("%.1f",sd(LE_Oceania$gdp_pc, na.rm = TRUE)), ")")
GDP_Europe <- paste0(
  sprintf("%.1f",mean(LE_Europe$gdp_pc, na.rm = TRUE)), ' (', sprintf("%.1f",sd(LE_Europe$gdp_pc, na.rm = TRUE)), ")")
GDP_NA <- paste0(
  sprintf("%.1f",mean(LE_NA$gdp_pc, na.rm = TRUE)), ' (', sprintf("%.1f",sd(LE_NA$gdp_pc, na.rm = TRUE)), ")")
GDP_SA <- paste0(
  sprintf("%.1f",mean(LE_SA$gdp_pc, na.rm = TRUE)), ' (', sprintf("%.1f",sd(LE_SA$gdp_pc, na.rm = TRUE)), ")")
GDP_All <- paste0(
  sprintf("%.1f",mean(LE_data$gdp_pc, na.rm = TRUE)), ' (', sprintf("%.1f",sd(LE_data$gdp_pc, na.rm = TRUE)), ")")

## Combine into vec
table1_GDP <- c(GDP_Africa, GDP_Asia, GDP_Oceania, GDP_Europe,
                GDP_NA, GDP_SA, GDP_All)

## Add ncountries
ncountries_Africa <- nrow(LE_Africa)
ncountries_Asia <- nrow(LE_Asia)
ncountries_Oceania <- nrow(LE_Oceania)
ncountries_Europe <- nrow(LE_Europe)
ncountries_NA <- nrow(LE_NA)
ncountries_SA <- nrow(LE_SA)
ncountries_All <- nrow(LE_data)

table1_ncountries <- c(ncountries_Africa, ncountries_Asia,
                       ncountries_Oceania, ncountries_Europe,
                       ncountries_NA, ncountries_SA, 
                       ncountries_All)

## Gather missingness % for each variable shown in table
life_expectancy_pmissing <- paste0(
  sprintf("%.1f",mean(is.na(LE_data$life_expectancy))*100), '%')

primary_completion_pmissing <- paste0(
  sprintf("%.1f",mean(is.na(LE_data$primary_completion_rate))*100), '%')

pop_density_pmissing <- paste0(
  sprintf("%.1f",mean(is.na(LE_data$pop_dens))*100), '%')

health_expenditure_pmissing <- paste0(
  sprintf("%.1f",mean(is.na(LE_data$health_expend_pc))*100), '%')

GDP_pmissing <- paste0(
  sprintf("%.1f",mean(is.na(LE_data$gdp_pc))*100), '%')

#Now combine all into df then table ----------------
table1_df <- c()
table1_df$Continent <- c("Africa", "Asia", "Australia/Oceania", "Europe", "North America", "South America", "Total")
table1_df$ncountries <- table1_ncountries
table1_df$Life_expectancy <- table1_life_expectancy
table1_df$primary_completion <- table1_primary_completion
table1_df$health_expenditure <- table1_health_expenditure
table1_df$pop_density <- table1_pop_density
table1_df$GDP <- table1_GDP

table1_df <- as.data.frame(table1_df)
names(table1_df) <- c("Continent",
                      "Number of Countries",
                      "Life Expectancy at Birth (years)",
                      "Primary School Completion Rate (total of % by age)", 
                      "Health Expenditure (current international $)",
                      "Population Density (people per sq. km)",
                      "GDP per Capita (current international $)")

table1_table <- flextable(table1_df, col_keys = c(
  "Continent",
  "Number of Countries",
  "Life Expectancy at Birth (years)",
  "Primary School Completion Rate (total of % by age)", 
  "Health Expenditure (current international $)",
  "Population Density (people per sq. km)",
  "GDP per Capita (current international $)"))

table1_table <- align(table1_table,j =1, align = "left", part = "all")
table1_table <- align(table1_table,j = 2:7, align = "right", part = "all")
table1_table <- width(table1_table, width = c(1.4,1.0,1.0,1.0,1.0,1.0, 1.0))
#table1_table <- height_all(table1_table,height = 0.30)
table1_table <- hrule(table1_table, rule = "at least", part = "all")
table1_table <- hline(table1_table, i = 6, j = NULL, part = "body", 
                      border = fp_border(color="black"))
#table1_table <- footnote(table1_table, j = 1, i = 1, part="body",  value = as_paragraph("Mean (SD)" ))
#table1_table <- footnote(table1_table, j = 1, i = 2, part="body",  value = as_paragraph(paste0(
#  "Value reported from complete cases only. Missingness present for each variable at the following rates: ",
#  "Life Expectancy ", life_expectancy_pmissing, ";",
#  "Primary Education Completion ", primary_completion_pmissing, ";",
#  "Health Expenditure ", health_expenditure_pmissing, ";",
#  "Population Density ", pop_density_pmissing, ";",
#  "GDP ", GDP_pmissing, ".")))

```

```{r, echo = FALSE}
table1_table
```

## Table of Variables with Missingness over 50%
```{r, echo = FALSE}
x <- c()

for (i in 1:ncol(LE_data))
{
  x[i] = mean(is.na(LE_data[,i]))
  names(x)[i] = names(LE_data)[i]
}

low_missing <- x[x < 0.5]
low_missing_vars = names(low_missing)
high_missing <- x[x >= 0.5]
high_missing_vars <- names(high_missing)
high_missing_vars_clean <- c("Children newly infected with HIV","Educational attainment, primary", "Educational attainment, bachelor's", "Literacy rate", "Poverty Ratio")

# Loop to get table vals
high_missing_char <- c()

for (i in 1:length(high_missing)){
  high_missing_char[i] <- paste0(sprintf("%.1f", high_missing[i]*100),
                             "%")
}

#Now combine all into df then table ----------------
table2_df <- c()
table2_df$Var <- high_missing_vars_clean
table2_df$nmissing <- high_missing_char

table2_df <- as.data.frame(table2_df)
names(table2_df) <- c("Variables",
                      "Missingness")

table2_table <- flextable(table2_df, col_keys = c("Variables",
                      "Missingness"))

table2_table <- align(table2_table,j =1, align = "left", part = "all")
table2_table <- align(table2_table,j = 2, align = "right", part = "all")
table2_table <- width(table2_table, width = c(1.4,1.4))
#table1_table <- height_all(table1_table,height = 0.30)
table2_table <- hrule(table2_table, rule = "at least", part = "all")
table2_table

# Note, we also manually added the variable with 100% missingness to the table after generation
```

## Some graphs for out variable of interest: LE
```{r, echo = FALSE}
p1 <- ggplot(LE_data, aes(life_expectancy)) +
  geom_histogram(color = "#000000", fill = "#0099F8") +
  theme_bw()+
  xlab("Life Expectancy (Years)")+
  ylab("Frequency")+ 
  theme(plot.margin = unit(c(1,0.5,0.5,0.5), "cm"))

## Plot 2: boxplot by GDP quantiles
gdp_quantiles <- quantile(LE_data$gdp_pc, c(0.25, 0.5, 0.75), na.rm = TRUE)
temp <- LE_data %>%
  dplyr::mutate(GDP_quantile = case_when(
    gdp_pc < gdp_quantiles[1] ~ "0-25%",
    gdp_pc < gdp_quantiles[2] ~ "25-50%",
    gdp_pc < gdp_quantiles[3] ~ "50-75%",
    gdp_pc >= gdp_quantiles[3] ~ "70-100%",
    
  )) %>%
  dplyr::filter(!(is.na(gdp_pc)))
p2 <- ggplot(temp, aes(x = GDP_quantile, y = life_expectancy, 
                 color = GDP_quantile)) +
  geom_boxplot(show.legend = FALSE) +
  ggsci::scale_color_jama()+
  theme(axis.text.x = element_text(angle = 45))+
  theme_bw()+
  xlab("GDP per Capita Percentile")+
  ylab("Life Expectancy (Years)") + 
  theme(plot.margin = unit(c(1,0.5,0.5,0.5), "cm"))

p3 <- ggplot(data = LE_data, aes(x = health_expend_pc, 
                                 y = life_expectancy))+
  geom_point(aes(size = pop_total, color = Continent),
             show.legend = c(TRUE))  + 
  ggsci::scale_color_jama()+
  scale_x_log10() + 
  theme_bw()+
  ylab("Life Expectancy (Years)") +
  xlab("Healthcare Expenditure Per Capita ($)") +
  guides(size = "none") +
  theme(plot.margin = unit(c(1,0.5,0.5,0.5), "cm"))

top <- cowplot::plot_grid(p1, p2, labels = c("A. Histogram of Life Expectancy", "B. Life Expectancy by GDP"),
                          hjust = -0.1)
bottom <- cowplot::plot_grid(p3, labels = c("C. Life Expectancy by Population and Healthcare Spending"),
                             hjust = -0.1)
cowplot::plot_grid(top, bottom, ncol = 1)


```

* The variable of primary interest is life expectancy. We show some graphs describing the life expectancy data for the 217 countries.

* We then select a subset of the 29 variables and showed the mean and standard deviation by continent. We selected one education, healthcare, financial, population

* We also observe that there is quite a bit of missingness here. We show the missingness rates for some variables here.

(b) Many predictors in the dataset contain missing values. Is deleting predictor variables with
many missing values an appropriate method to deal with missing values? Choose a method
to deal with the missing values and then employ this method to the life expectancy data.
Justify your choice. Additionally, there are some countries (cases) where the value of Life
expectancy is missing. Explain how you will handle this problem. [13 marks]

```{r, echo = FALSE}

## Subset variables with missingness < 50%
LE_data1 <- LE_data[ ,names(LE_data) %in% low_missing_vars]
Country.Name <- LE_data$Country.Name

LE_data2 <- subset(LE_data1, select = -c(Country.Name)) 
# remove ID column for imputation

```

```{r, echo = FALSE}

#cor(LE_data2, use = "complete.obs")
#install.packages("caret")
#install.packages("vctrs")
#library(caret)
#findLinearCombos(LE_data2)

```

```{r, echo = FALSE, results = 'hide'}
### First, we create predictor matrix without LE, so LE isn't used
### to predict covariates.

pred <- mice::make.predictorMatrix(LE_data2)
pred[,"life_expectancy"] <- 0 # never use it as a predictor



imputations <- mice(LE_data2, seed = 231, 
                    predictorMatrix = pred,
                    method = "pmm", m = 10)

imputations$imp
#stripplot(imputations, pch = 20, cex = 1.2)
#xyplot(imputations, bmi ~ chl | .imp, pch = 20, cex = 1.4)

#Define the analysis you want to perform at Stage 2
feature.selection1 <- expression(null.model1 <- lm(life_expectancy ~ 1), model2 <- step(null.model1, scope = ~ Continent + electricity_access + income_growth + income_growth_pc + children_out_of_school + infant_mortality_rate + primary_completion_rate + real_interest_rate + pop_growth_perc_annual + pop_dens + pop_total + health_expend_pc + health_expend_perc_gdp + unemployment_rate + gdp_growth + gdp_pc + birth_rate + adults_HIV + safe_drinking_water + ed_duration ))

feature.selection1 <- expression(null.model1 <- lm(life_expectancy ~ 1), model2 <- step(null.model1, scope = ~ Continent + electricity_access + income_growth + income_growth_pc + children_out_of_school + infant_mortality_rate + primary_completion_rate + real_interest_rate + pop_growth_perc_annual + pop_dens + pop_total + health_expend_pc + health_expend_perc_gdp + unemployment_rate + gdp_growth + gdp_pc + birth_rate + adults_HIV + safe_drinking_water + ed_duration ))

#Using the with function evaluate the above expression
step.fit <- with(imputations, feature.selection1)
step.fit.models <- lapply(step.fit$analyses, formula)
step.fit.features <- lapply(step.fit.models, terms)
feature.frequency <- unlist(lapply(step.fit.features, labels))
feature.frequency
sort(table(feature.frequency), decreasing=TRUE)

```

```{r, echo = FALSE}
# Plot the missing values manually
countries_missing_LE <- dplyr::filter(LE_data, is.na(life_expectancy))$Country.Name

vars <- c("life_expectancy", "adults_HIV", "health_expend_pc",
          "infant_mortality_rate")
imp1 <- cbind(LE_data$Country.Name,"imp" = 1, 
              complete(imputations, 1))
imp2 <- cbind(LE_data$Country.Name,"imp" = 2, 
              complete(imputations, 2))
imp3 <- cbind(LE_data$Country.Name,"imp" = 3, 
              complete(imputations, 3))
imp4 <- cbind(LE_data$Country.Name,"imp" = 4, 
              complete(imputations, 4))
imp5 <- cbind(LE_data$Country.Name,"imp" = 5, 
              complete(imputations, 5))

imps <- rbind(imp1, imp2, imp3, imp4, imp5)
names(imps)[1] <- "Country.Name"
imps <- imps %>%
  dplyr::mutate(Type = ifelse(Country.Name %in% countries_missing_LE, "Imputed", "Not Imputed")) %>%
  dplyr::arrange(desc(Type)) # so 

p1 <- ggplot(imps, aes(x = imp, 
                       y = life_expectancy,
                       color = Type)) +
  geom_point(size = 0.8, shape = 1,
             show.legend = FALSE) + 
  theme_bw()+ 
  scale_color_manual(values = c("#8B0000", "#FFFF00")) + 
  xlab("Imputation") + 
  ylab("Life expectancy (years)")

p2 <- ggplot(imps, aes(x = imp, 
                       y = safe_drinking_water,
                       color = Type)) +
  geom_point(shape = 1, size = 0.8, show.legend = FALSE) +
  theme_bw()+ 
  scale_color_manual(values = c("#8B0000", "#FFFF00")) + 
  xlab("Imputation") + 
  ylab("Safe drinking water (%)")

p3 <- ggplot(imps, aes(x = imp, 
                       y = health_expend_pc,
                       color = Type)) +
  geom_point(shape = 1, size = 0.8, show.legend = FALSE) +
  theme_bw()+ 
  scale_color_manual(values = c("#8B0000", "#FFFF00")) + 
  xlab("Imputation") + 
  ylab("Health expenditure per capita ")

p4 <- ggplot(imps, aes(x = imp, 
                       y = infant_mortality_rate,
                       color = Type)) +
  geom_point(shape = 1, size = 0.8, show.legend = FALSE) +
  theme(axis.title.y = element_text(size = 9)) + 
  theme_bw()+ 
  scale_color_manual(values = c("#8B0000", "#FFFF00")) + 
  xlab("Imputation") + 
  ylab("Infant mortality rate (per 1,000 live births)")

cowplot::plot_grid(p1, p2, p3, p4, nrow = 2, hjust = -0.1)
```

2. Collinearity increases the variance of the estimators and hence, reduces the adequacy of the model. When collinearity is present, how do you solve this problem? Investigate collinearity between the predictor variables in the LifeExpectancyData1.csv dataset. [12 marks]

```{r, echo = FALSE}
# Use faraway from Lab 4
#model1 <- lm(life_expectancy ~ ., data=complete(imputations))
#summary(model1)

#corr <- Hmisc::rcorr(as.matrix(LE_data2[,-1]))
#corr

#corrplot.mixed(corr, lower.col = "black", number.cex = 0.7, tl.cex = 0.5)

#corrplot(corr,tl.cex = 0.5)

#faraway::vif(complete(imputations)[,-1])
```

```{r, echo = FALSE}
#####
### Take 2 with miceadds package
#####

# calculate correlations. Remove Continent (factor) and life expectancy (response)
mi_corr <- micombine.cor(imputations, variables = c(3:21))

# This gives pairwise correlations in columns.
# turn it into a matrix (for plotting, analysis)
cor_df <- mi_corr %>%
  dplyr::select(variable1, variable2, r) %>%
  tidyr::pivot_wider(names_from = variable2, values_from = r)

# rearrange cols, make diagonals 1 instead of NA
cor_df <- cor_df %>%
  dplyr::relocate(electricity_access, .before = income_growth)  %>%
  dplyr::mutate_all(~replace(., is.na(.), 1))

# lastly move var names from col 1 into row names
cor_df <- as.data.frame(cor_df)
rownames(cor_df) <- cor_df$variable1
cor_df <- cor_df[,-1]

## Set names for corplot printing after this
colnames(cor_df) <-  c("Electricity access", "Income growth",
                       "Income growth percent", "Children out of school",
                       "Infant mortality rate", "Primary completion rate",
                       "Real interest rate", "Percent pop. growth",
                       "Pop. density", "Total pop.",
                       "Health spending per capita",
                       "Health spending perc. GDP",
                       "Unemployment rate",
                       "GDP growth",
                       "GDP per capita",
                       "Birth rate",
                       "Adults HIV perc.",
                       "Safe drinking water",
                       "Education duration")
rownames(cor_df) <- colnames(cor_df)
  
# Plot
#corrplot.mixed(as.matrix(cor_df), lower.col = "black", number.cex = 0.7, tl.cex = 0.5)
corrplot(as.matrix(cor_df),tl.cex = 0.8, tl.col="black")

# VIF. Find mean over imputations initially
vif_output <- list()
for (i in 1:10){
  vif_output[[i]] <- faraway::vif(complete(imputations, i)[,-c(1,2)])
}
vif_output_means <- c()
for (i in 1:length(vif_output[[1]])){
  val_vec <- c()
  for (j in 1:10){
    val_vec[j] <- vif_output[[j]][i]
  }
  vif_output_means[i] <- mean(val_vec)
  names(vif_output_means)[i] <- names(vif_output[[1]][i])
}
vif_output_means

# VIF. Find mean over imputations after
drop_vars <- c("Continent","life_expectancy","income_growth","birth_rate", "electricity_access", "health_expend_pc")


vif_output_after <- list()
for (i in 1:10){
  vif_output_after[[i]] <- faraway::vif(complete(imputations, i)[,!(names(complete(imputations, i)) %in% drop_vars)])
}
vif_output_means_after <- c()
for (i in 1:length(vif_output_after[[1]])){
  val_vec_after <- c()
  for (j in 1:10){
    val_vec_after[j] <- vif_output_after[[j]][i]
  }
  vif_output_means_after[i] <- mean(val_vec_after)
  names(vif_output_means_after)[i] <- names(vif_output_after[[1]][i])
}
vif_output_means_after
```

```{r}
### VIF table
temp <- data.frame("Feature" = names(vif_output_means_after),
                   "VIF After Dropping Vars" = vif_output_means_after)

tableS1_df <- data.frame("Feature" = colnames(cor_df),
                         "VIF" = sprintf("%.2f",vif_output_means))
tableS1_df <- dplyr::left_join(tableS1_df, temp, by = "Feature")%>%
  dplyr::mutate(VIF.After.Dropping.Vars = 
                  ifelse(is.na(VIF.After.Dropping.Vars),
                         " ",
                         sprintf("%.2f", VIF.After.Dropping.Vars)))

names(tableS1_df)[3] <- "VIF After Dropping Vars"


tableS1_table <- flextable(tableS1_df)

tableS1_table <- align(tableS1_table,j =1, align = "left", part = "all")
tableS1_table <- align(tableS1_table,j =2:3, align = "right", part = "all")
tableS1_table <- width(tableS1_table, width = c(1.4,1.0, 1.0))
#table1_table <- height_all(table1_table,height = 0.30)
tableS1_table <- hrule(tableS1_table, rule = "at least", part = "all")
tableS1_table


```

We see a number of variables that have high VIFs and correlations with similar values with absolute values > 0.8. Based on these observations, we decide to remove national net income growth because of its high correlation with national net income growth per capita (r = 0.979), birth rate because of its high positive correlation with population growth rate (r = 0.76), infant mortality rate (r = 0.86) and large negative correlation with safe drinking water rate (r = - 0.85), and healthcare expenditure per capita because of its high correlation with GDP per capita (r = 0.81) and healthcare expenditure as percent of GDP (r = 0.51). We also decide to remove electricity access because of its high VIF and large correlations with infant mortality rate (r = 0.81), primary completion rate (r = 0.71), and safe drinking water rate (r = 0.72). Removing these helps us address multicollinearity, as evidenced by the larrge drop in VIF (Table S1), which makes our coefficients and their standard errors more appropriate. 

3. To understand better life expectancy and the factors that affect it, suggest the best linear model which predicts life expectancy in 2020. Interpret and evaluate the suggested model. [25 marks]

```{r, echo = FALSE, results = 'hide'}
stripplot(imputations, pch = 20, cex = 1.2)
#xyplot(imputations, bmi ~ chl | .imp, pch = 20, cex = 1.4)

#Define the analysis you want to perform at Stage 2
feature.selection1 <- expression(null.model1 <- lm(life_expectancy ~ 1), model2 <- step(null.model1, scope = ~  income_growth_pc + children_out_of_school + infant_mortality_rate + primary_completion_rate + real_interest_rate + pop_growth_perc_annual + pop_dens + pop_total + health_expend_perc_gdp + unemployment_rate + gdp_growth + gdp_pc + adults_HIV + safe_drinking_water + ed_duration, direction = "forward"))

#feature.selection1 <- expression(null.model1 <- lm(life_expectancy ~ 1), model2 <- step(null.model1, scope = ~  income_growth_pc + infant_mortality_rate + pop_growth_perc_annual + pop_dens + pop_total + health_expend_pc + health_expend_perc_gdp + gdp_growth + gdp_pc +  ed_duration))

#Using the with function evaluate the above expression
step.fit <- with(imputations, feature.selection1)
step.fit.models <- lapply(step.fit$analyses, formula)
step.fit.features <- lapply(step.fit.models, terms)
feature.frequency <- unlist(lapply(step.fit.features, labels))
feature.frequency
```

Stepwise feature selection with our imputed datasets yields the following table showing the frequency of feature selection:

```{r, echo = FALSE}
sort(table(feature.frequency), decreasing=TRUE)
```

We observe that five features appear in at least 80% of the final models after stepwise selection, so we decide to build a model with only these five variables.

```{r, echo = FALSE}
# Ge
fit <- with(imputations, lm(life_expectancy ~ 
         health_expend_perc_gdp + infant_mortality_rate + safe_drinking_water + 
         gdp_pc))

fit$analyses %>% str(max.level = 1)
  
pooled_results <- mice::pool(fit)
summary(pooled_results)

```

```{r, echo = FALSE}
# Create table showing output
m <- 10
df <- as.data.frame(summary(pooled_results))
table3_df <- c()
table3_df$Feature <- c("Intercept",
                        "Number of adults with HIV", 
                       "Infant mortality rate",
                       "% drinking safe water",
                       "GDP per capita")
table3_df$Estimate <- sprintf("%.4f", df$estimate)
table3_df$SE <- sprintf("%.5f", df$std.error)
table3_df$df <- sprintf("%.2f", df$df)
table3_df$P_val <- sprintf("%.4f", df$p.value)
table3_df$P_val <- ifelse(table3_df$P_val == "0.0000",
                          "< 0.0001", table3_df$P_val)
table3_df$BestSubsetFrequency <- c(" ",
  table(feature.frequency)[names(table(feature.frequency)) == "adults_HIV"]/m,
  table(feature.frequency)[names(table(feature.frequency)) == "infant_mortality_rate"]/m,
  table(feature.frequency)[names(table(feature.frequency)) == "safe_drinking_water"]/m,
  table(feature.frequency)[names(table(feature.frequency)) == "gdp_pc"]/m
)

# recalculate VIF for these five vars only
vif_output <- list()
for (i in 1:10){
  vif_output[[i]] <- faraway::vif(dplyr::select(
    complete(imputations, i), adults_HIV, infant_mortality_rate,
     safe_drinking_water, gdp_pc))
}
vif_output_means2 <- c()
for (i in 1:length(vif_output[[1]])){
  val_vec <- c()
  for (j in 1:10){
    val_vec[j] <- vif_output[[j]][i]
  }
  vif_output_means2[i] <- mean(val_vec)
  names(vif_output_means2)[i] <- names(vif_output[[1]][i])
}
vif_output_means2
table3_df$VIF <- c(" ",vif_output_means2)
table3_df <- as.data.frame(table3_df)


names(table3_df) <- c("Feature",
                      "Estimate",
                      "Standard Error",
                      "df",
                      "P-value",
                      "Proportion of Best Subsets Appeared in",
                      "VIF")

table3_table <- flextable(table3_df, col_keys = c("Feature",
                      "Estimate",
                      "Standard Error",
                      "df",
                      "P-value",
                      "Proportion of Best Subsets Appeared in",
                      "VIF"))

table3_table <- align(table3_table,j =1, align = "left", part = "all")
table3_table <- align(table3_table,j =2:7, align = "right", part = "all")
table3_table <- width(table3_table, width = c(1.4,1.0, 1.1, 1.3, 1.0, 1.0, 1.0 ))
#table1_table <- height_all(table1_table,height = 0.30)
table3_table <- hrule(table3_table, rule = "at least", part = "all")
table3_table <- vline(table3_table, i = NULL, j = 5, part = "body", 
                      border = fp_border(color="black"))
table3_table
```


```{r, echo = FALSE}

# 
predict_df <- cbind(
  predict(fit$analyses[[1]], newdata = complete(imputations, 1)),
  predict(fit$analyses[[2]], newdata = complete(imputations, 2)),
  predict(fit$analyses[[3]], newdata = complete(imputations, 3)),
  predict(fit$analyses[[4]], newdata = complete(imputations, 4)),
  predict(fit$analyses[[5]], newdata = complete(imputations, 5)),
  predict(fit$analyses[[6]], newdata = complete(imputations, 6)),
  predict(fit$analyses[[7]], newdata = complete(imputations, 7)),
  predict(fit$analyses[[8]], newdata = complete(imputations, 8)),
  predict(fit$analyses[[9]], newdata = complete(imputations, 9)),
  predict(fit$analyses[[10]], newdata = complete(imputations, 10))
)

observed_df <- cbind(
  dplyr::select(complete(imputations, 1), life_expectancy),
  dplyr::select(complete(imputations, 2), life_expectancy),
  dplyr::select(complete(imputations, 3), life_expectancy),
  dplyr::select(complete(imputations, 4), life_expectancy),
  dplyr::select(complete(imputations, 5), life_expectancy),
  dplyr::select(complete(imputations, 6), life_expectancy),
  dplyr::select(complete(imputations, 7), life_expectancy),
  dplyr::select(complete(imputations, 8), life_expectancy),
  dplyr::select(complete(imputations, 9), life_expectancy),
  dplyr::select(complete(imputations, 10), life_expectancy)
)

residual_df <- predict_df - observed_df
names(residual_df) <- c(1:10)

# Residual_df has 15 columns (one for each imp) and 217  rows
# (one per obs). Now we find the mean for each obs and use that
# for residual diagnostic plots

residual_means <- residual_df %>%
  dplyr::mutate(mean =rowMeans(residual_df))
residual_means <- residual_means[,"mean"]

## And do same for fit (for residual vs fit)
predicted_means <- as.data.frame(predict_df) %>%
  dplyr::mutate(mean = rowMeans(predict_df))
predicted_means <- predicted_means[,"mean"]

mean_df <- as.data.frame(cbind(predicted_means, residual_means))

####
## Now plot
####

p1 <- ggplot(mean_df, aes(x=residual_means))+
  geom_histogram() + 
  ggsci::scale_colour_jama() + 
  theme_bw() + 
  theme(plot.margin = unit(c(1,0.5,0.5,0.5), "cm")) + 
  xlab("Residuals") + 
  ylab("Count")
p2 <- ggplot(mean_df, aes(sample = residual_means)) +
  stat_qq() +
  stat_qq_line() + 
  theme_bw()+ 
  theme(plot.margin = unit(c(1,0.5,0.5,0.5), "cm")) +
  xlab("Theoretical") +
  ylab("Sample")
p3 <- ggplot(mean_df, aes(x=predicted_means, y = residual_means))+
  geom_point() + 
  theme_bw() + 
  theme(plot.margin = unit(c(1,0.5,0.5,0.5), "cm")) +
  xlab("Fitted Value")+
  ylab("Residual")

top <- plot_grid(p1, p2, labels = c("A. Histogram of Residuals",
                                    "B. QQplot of Residuals"),
                 hjust = -0.1)
bottom <- plot_grid(p3, labels = "C. Residuals vs. Fit",
                    hjust = -0.1)
plot_grid(top, bottom, nrow = 2)



```

The figure above shows evidence that the residuals produced by our model predictions are approximately normally distributed without a clear pattern. This evidence is a sign that our model is good.  


4. Using the same dataset and using the additional feature Continent, employ an appropriate experimental design to study differences of average life expectancies across the continents: Asia, Europe, North America, South America, Africa, Australia/Oceania. Justify your choice of experimental design and methods. [13 marks]


```{r, echo = FALSE}
library(car)
anova_LE <- aov(life_expectancy ~ as.factor(Continent), data=complete(imputations))
summary(anova_LE)
Anova(anova_LE, type = "III")

```
To study the differences in life expectancy across continents (a factor variable), we employ an ANCOVA experimental design. ANCOVA is designed specifically for testing whether there is a difference in means for a continuous response variable (in our case, life expectancy) across a categorical predictoor (continent), and it has the additional benefit of controlling for covariates that are independent of the predictor. In our experimental design, we perform an ANCOVA with % of GDP spent on health as a covariate. 

Due to the fact that the references and software for performing ANCOVA and related post-hoc analyses on multiply imputed data are limited, we use single imputation by chained equations (multiple imputation with m = 1) for this experiment. To combat the limitations of only using one imputation, we will also perform sensitivity analyses with different random seeds given to the imputation algorithm to ensure that was not do to the randomness of the single imputation.

```{r, echo = FALSE}
# We can just pick complete(imputations, 1) as our single imputation
anova_data <- complete(imputations, 1)
anova_data$Continent <- as.factor(anova_data$Continent)

# perform ANCOVA (according to slides from Lecture 10)
le.aov <- aov(life_expectancy ~ log(gdp_pc) + Continent,
               data = anova_data)
summary(le.aov)

# post-hoc
tukey.le <- glht(le.aov, linfct = mcp(Continent = "Tukey"))

#view a summary of the post hoc comparisons
summary(tukey.le)
```


Before proceeding with the statistical tests, we check the assumptions of ANCOVA which are (potentially show diagnostic plots in appendix only):

* Homogeneity of regression slopes.
* The relationship between the response and the covariate is linear.
```{r, echo = FALSE}
# Plot life expectancy by % of GDP spent on healthcare
p1 <- ggplot(anova_data, aes(x = log(gdp_pc), 
                             y = life_expectancy, 
                             color = Continent)) +
  theme_bw()+
  geom_point(alpha = 0.5, size = 0.8)+
  xlab("log(GDP per capita in thousands)")+
  ylab("Life expectancy") +
  ggsci::scale_color_jama()+
  geom_smooth(method="lm",se = FALSE)

# Figure showing life expectancy by continent (with singly imputed dataset that we will use)

p2 <- ggplot(complete(imputations, 1), aes(x = Continent,
                                        y = life_expectancy, 
                 color = Continent)) +
  geom_boxplot(show.legend = FALSE) +
  ggsci::scale_color_jama() +
  theme_bw()+
  theme(axis.text.x = element_text(angle = -45),
        plot.margin = unit(c(1,0.5,0.5,0.5), "cm")) +
  xlab("Continent")+
  ylab("Life Expectancy (Years)") 

plot_grid(p1, p2, labels = c("A", "B"))

```

```{r, echo = FALSE}
# test for homogeneity of variances
car::leveneTest(anova_data$life_expectancy~anova_data$Continent)
```

* Normality of residuals
* Equal variances between the different continents

```{r}
# Look at residuals
anova_data$residuals <- le.aov$residuals
p1 <- ggplot(anova_data, aes(residuals)) +
  geom_histogram(color = "#000000", fill = "#0099F8") +
  theme_bw()+
  xlab("Residual")+
  ylab("Frequency")+ 
  theme(plot.margin = unit(c(1,0.5,0.5,0.5), "cm"))
p2 <- ggplot(anova_data, aes(sample = residuals)) +
  stat_qq() +
  stat_qq_line() + 
  theme_bw()+ 
  theme(plot.margin = unit(c(1,0.5,0.5,0.5), "cm")) +
  xlab("Theoretical") +
  ylab("Sample")

plot_grid(p1, p2, labels = c("A.", "B"))

shapiro.test(anova_data$residuals1)
```



(Null hypothesis and alternative hypothesis here)